Using Component Model to implement a DSL: The Multi-Stencil Language Case Study

### 2016-11-04 ARTICLES/OTHER/C2S_EXA/slides.tex
WIP ..

%-------------------------------------------------------------------
\title[The Multi-Stencil Language]{Using Component Model to implement a DSL: The Multi-Stencil Language Case Study}
\author[]{Hélène Coullon, Julien Bigot, \underline{Christian Perez}} % [Hélène Coullon (INRIA), Julien Bigot (CEA), Christian Perez (INRIA)]
\institute[]{Avalon project team\\Maison de la simulation}
\date{C2S@Exa -- Scientific Days \\November 8th, 2016\\ Paris}
%-------------------------------------------------------------------

@@

\begin{frame}
    \titlepage
\end{frame}

@@

%-------------------------------------------------------------
\section{From DSL to Component Assembly}
%-------------------------------------------------------------
%-------------------------------------------------------------
\begin{frame}
\frametitle{What is produced ?}
A data parallel/task parallel component assembly.\vspace{-0.2em}

@@

%----------------
\begin{frame}
\frametitle{Data and Task parallelism}
\begin{block}{Synchronizations}
When a computation read a data, using a stencil shape, that has been written by a previous computation.
\end{block}

\begin{center}
\textcolor{red}{Static} or \textcolor{red}{dynamic} scheduling ?
\end{center}
\end{frame}
%----------------
%% \begin{frame}
%% \frametitle{Series-Parallel Tree}
%% \small\textit{Valdes \& Al, The Recognition of Series Parallel Digraphs, STOC '79}
%% \normalsize
%% \begin{columns}
%% \begin{column}{.48\textwidth}
%% \begin{center}
%% \resizebox{\textwidth}{!}{%
%% \begin{tikzpicture}[shorten >=1pt, node distance=2cm, on grid, auto]
%%    \node[] (s0)    at (0,0) {$\mathcal{S}$};

@@

%-------------------------------------------------------------
\section{Evaluations}
%----------------
\begin{frame}
\frametitle{Evaluations}
\begin{block}{Application}
\begin{itemize}
\item FullSWOF2D: developed at the MAPMO, Université d'Orléans
\item Solve the shallow-water equations using a finite volume method
\item Cartesian mesh
\item 3 mesh entities, 7 computation domains, 48 quantities
\item 98 computations (32 stencils, 66 local kernels)
\end{itemize}
\end{block}
\end{frame}
%----------------

\begin{frame}
\frametitle{Evaluations}
\begin{block}{Evaluations}
\begin{itemize}
\item Data parallelization: no overhead
  \begin{itemize}
  \item weak scaling, fix size 400x400 (16.384 cores)
  \item strong scaling, 10kx10k:1k (16.384 cores)
  \end{itemize}
\item Hybrid parallelization: increase performance\\
  \begin{itemize}
  \item when data parallelism reaches its limits
  \item introduce new parallelism
  \end{itemize}
\end{itemize}
\end{block}

\begin{block}{Machines}
\begin{itemize}
\item Thin nodes TGCC Curie
\item 2 CPU, 8-cores, Intel Sandy Bridge EP (E5-2680) 2.7 GHz, 64 Go
\item OpenMPI and gcc 4.9
\end{itemize}
\end{block}
\end{frame}
%-------------------------------------------------------------
\begin{frame}
\frametitle{Evaluations}
\begin{columns}
\begin{column}{6cm}
\resizebox{6cm}{!}{\includegraphics{images/median_weak.pdf}}
\end{column}
\begin{column}{6cm}
\resizebox{6cm}{!}{\includegraphics{images/median_strong.pdf}}
\end{column}
\end{columns}
\begin{center}
\textbf{No overhead introduced by components}
\end{center}
\end{frame}
%-------------------------------------------------------------
\begin{frame}
\frametitle{Evaluations}
\begin{columns}
\begin{column}{6cm}
\resizebox{6cm}{!}{\includegraphics{images/times.pdf}}
\end{column}
\begin{column}{6cm}
\resizebox{6cm}{!}{\includegraphics{images/base_close_median.pdf}}
\end{column}
\end{columns}
\begin{center}
\textbf{Good performance with 8 threads}
\end{center}
\end{frame}
%-------------------------------------------------------------

%----------------
\begin{frame}
\frametitle{Evaluations}
\textbf{Improvment with data+task scheduling}
\begin{center}
\resizebox{8cm}{!}{\includegraphics{./images/base_close_median.pdf}}
\end{center}
\end{frame}

%-------------------------------------------------------------
\section{Conclusion}
%-------------------------------------------------------------
%\subsection{Conclusion and perspectives}
\begin{frame}
\frametitle{Conclusion}
\begin{block}{Conclusion}
\begin{itemize}
\item A DSL for Multi-Stencil applications (MSL)
\item The compilation of MSL to get a parallel scheduling pattern of the simulation
\begin{itemize}
\item Data parallelism
\item Task parallelism
\end{itemize}
\item The dump to a component-based runtime
\item Data parallelism evaluation: no overhead introduced 
\end{itemize}
\end{block}
\end{frame}
%-------------------------------------------------------------

%-------------------------------------------------------------
\begin{frame}
\begin{block}{Perspectives}
\frametitle{Perspectives}
\begin{itemize}
\item Improvment of the language (convergence criteria, reduction etc.)
\item Scalability up to $32k$ cores on TGCC Curie (CEA)
\item Evaluations on Data+Task parallelism
\begin{itemize}
\item OpenMP 3 inside kernels
\end{itemize}
\item Dynamic scheduling
\begin{itemize}
\item OpenMP 4 with a scheduling component
\item Kstar for StarPU and XKaapi runtimes
\end{itemize}
\item CPU+GPGPUs using stencil compilers (Pochoir, PATUS etc.)
\end{itemize}
$\hookrightarrow$ Show portability, maintainability introduced by components
\end{block}
\end{frame}
%-------------------------------------------------------------

>>>>

%-------------------------------------------------------------------
\title[The Multi-Stencil Language]{Using Component Model to implement a DSL: The Multi-Stencil Language Case Study}
\author[]{Hélène Coullon, Julien Bigot, \underline{Christian Perez}} % [Hélène Coullon (INRIA), Julien Bigot (CEA), Christian Perez (INRIA)]
\institute[]{Avalon project team\\Maison de la simulation}
\date{C2S@Exa -- Scientific Days \\Paris, November 8th, 2016}
%-------------------------------------------------------------------

@@

\begin{frame}
    \titlepage
    \vspace{1.25cm}
    \resizebox{2.5cm}{!}{\includegraphics{images/logo-inria.png}}\hspace{1cm}
    \resizebox{1cm}{!}{\includegraphics{images/logo-mdls.png}}\hfill
    \resizebox{1cm}{!}{\includegraphics{images/elci_logo.png}}
\end{frame}

@@

%-------------------------------------------------------------
\section{From DSL to Component Assembly}
%-------------------------------------------------------------
%----------------
\begin{frame}
\frametitle{Data and Task parallelism}
\begin{block}{Synchronizations}
When a computation read a data, using a stencil shape, that has been written by a previous computation.
\end{block}

\begin{center}
\textcolor{red}{Static} or \textcolor{red}{dynamic} scheduling ?
\end{center}
\end{frame}
%----------------
%----------------
\begin{frame}
\frametitle{MSL to Component-based runtime}
\begin{block}{Ready-to-fill parallel pattern}
\begin{itemize}
\item Data parallelism
\begin{itemize}
\item External distributed data structure
\item Automatic detection of synchronizations
\end{itemize}
\item Task parallelism (mid-grain)
\begin{itemize}
\item Static scheduling at the assembly level 
\item Dynamic scheduling within a driver component
  \begin{itemize}
  \item OpenMP task (For loop also available)
  \end{itemize}
\end{itemize}
\end{itemize}
\end{block}
The fine grain task parallelism is left to other languages: 
\begin{itemize}
\item OpenMP in the kernels
\item Kernels generated by stencil compilers
  \begin{itemize}
  \item Pochoir, PATUS, Liszt etc.
  \end{itemize}
\end{itemize}
\end{frame}
%----------------
%-------------------------------------------------------------
\begin{frame}
\frametitle{Component Assembly Runtime Template}
A data parallel/task parallel component assembly.\vspace{-0.2em}

@@

%% \begin{frame}
%% \frametitle{Series-Parallel Tree}
%% \small\textit{Valdes \& Al, The Recognition of Series Parallel Digraphs, STOC '79}
%% \normalsize
%% \begin{columns}
%% \begin{column}{.48\textwidth}
%% \begin{center}
%% \resizebox{\textwidth}{!}{%
%% \begin{tikzpicture}[shorten >=1pt, node distance=2cm, on grid, auto]
%%    \node[] (s0)    at (0,0) {$\mathcal{S}$};

@@

%-------------------------------------------------------------
\section{Evaluations}
%----------------
\begin{frame}
\frametitle{Evaluations}
\begin{block}{Application}
\begin{itemize}
\item FullSWOF2D: developed at the MAPMO, Université d'Orléans
\item Solve the shallow-water equations using a finite volume method
\item Cartesian mesh
\item 3 mesh entities, 7 computation domains, 48 quantities
\item 98 computations (32 stencils, 66 local kernels)
\end{itemize}
\end{block}
\begin{block}{Machines}
\begin{itemize}
\item Thin nodes TGCC Curie
\item 2 CPU, 8 cores, Sandy Bridge EP (E5-2680) 2.7 GHz, 64 GB
\item OpenMPI and gcc 4.9
\end{itemize}
\end{block}
\end{frame}
%----------------
%-------------------------------------------------------------
\begin{frame}
\frametitle{Data Parallelization: Weak and Strong Scaling}
\begin{columns}
\begin{column}{6cm}
\resizebox{6cm}{!}{\includegraphics{images/median_weak.pdf}}
\begin{center}
\vspace{-1em}
  Weak Scaling\\
  400x400 domain/core
\end{center}
\end{column}
\begin{column}{6cm}
\resizebox{6cm}{!}{\includegraphics{images/median_strong.pdf}}
\begin{center}
\vspace{-1em}
  Strong Scaling\\
  10k x 10k domaine
\end{center}
\end{column}
\end{columns}
\bigskip
\begin{center}
  \textbf{No overhead introduced by components}
\end{center}
\end{frame}
%-------------------------------------------------------------
\begin{frame}
\frametitle{Data Paralization to Hybrid Parallization: Strong Scaling}
\begin{columns}
\begin{column}{6cm}
  \resizebox{6cm}{!}{\includegraphics{images/times.pdf}}
  \begin{center}
\vspace{-1em}\small    Data Parallelization    
  \end{center}
\end{column}
\begin{column}{6cm}
\resizebox{6cm}{!}{\includegraphics{images/base_close_median.pdf}}
\begin{center}
\vspace{-1em}\small  Data and Hybrid Parallelizations
\end{center}
\end{column}
\end{columns}
\bigskip
\begin{center}
{\small
  \begin{tabular}{|c|c|c|c|c|c|c|c|c|}
    \hline 
   Parallelism Level & 1 & 2 & 3 & 4 & 6 & 10 & 12 & 16\\
   \hline
   Frequency & 2 & 1 & 3 & 5 & 3 & 1 & 1 & 2\\
   \hline
 \end{tabular}
 \\
 Task graph analysis
}
\end{center}
\end{frame}
%-------------------------------------------------------------

%-------------------------------------------------------------
\section{Conclusion}
%-------------------------------------------------------------
%\subsection{Conclusion and perspectives}
\begin{frame}
\frametitle{Conclusion and Future Work}
\begin{block}{Conclusion}
\begin{itemize}
\item A DSL for Multi-Stencil applications (MSL)
\item The generation of a component based runtime, including scheduling
\begin{itemize}
\item Data parallelism
\item Task parallelism
\end{itemize}
\item Evaluation on Curie: no overhead introduced 
\end{itemize}
\end{block}

\begin{block}{Future Work}
  \begin{itemize}
  \item Language improvement (convergence criteria, reduction etc.)
  \item OpenMP 3 inside kernels
  \item CPU+GPGPUs using stencil compilers (Pochoir, PATUS etc.)
$\hookrightarrow$ Show portability, maintainability introduced by components

  \end{itemize}
\end{block}
\end{frame}
%-------------------------------------------------------------
